---
title: "Reference_Genes_qPCR"
author: "Anna Duenser"
date: "`r format(Sys.time(), '%d %B, %Y')`"
#github-repo: "https://github.com/"
output:
  bookdown::gitbook:
    config:
      fontsettings:
        theme: night
        family: sans
        size: 2
  bookdown::pdf_document2:
    keep_tex: true
header-includes:
- \usepackage{hyperref}
- \usepackage{float}
- \usepackage{caption}
- \captionsetup[figure]{font=small}
#subtitle: University of Graz, Institut for Biology
#bibliography: bib.bibtex
#biblio-style: authoryear
#link-citations: yes
urlcolor: blue
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  fig.pos   = "H",
  out.extra = "",
  fig.align = "center",
  out.width = "100%",
  echo      = TRUE,
  include   = TRUE
  )
# Our Libraries
VLIBS <- c("tidyverse", "patchwork", "here", "readxl", "bookdown")
# Use Colors in Plots
# https://stackoverflow.com/questions/42458412/plotting-data-by-color-variable-with-ggplot2-in-r#comment72062654_42458412
colorBlindGrey  <- c("#C5C1C1" ,"#464343", "#E69F00", "#56B4E9", "#009E73",
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
F_LoadLibs <- function(x){
  print(x)
  if(!require(x, character.only = TRUE)) install.packages(x); library(x, character.only = TRUE);
  return(T)
}
# https://stackoverflow.com/questions/62140483/how-to-interpret-dplyr-message-summarise-regrouping-output-by-x-override
options(dplyr.summarise.inform=F)
# https://stackoverflow.com/questions/13286531/how-to-suppress-warnings-when-plotting-with-ggplot
options(warn=-1)
sapply(VLIBS, F_LoadLibs)
rm(VLIBS, F_LoadLibs)

```


"To choose stably expressed reference genes, we ranked the genes with a small log2 foldchange (-0.05 > l2fc < 0.05) for each separate gene expression analysis. After that we computed the coefficient of variation (CV) for each gene, across the samples for eahc lake of the normalized count matrices. We then checke which genes overlap between the two lakes and added the two CV numbers to order for the lowest CV change.  Ten reference genes were then selected from this list (including 38 genes) based on their base mean expression level to perform qPCR analysis."


```{r}
library("tidyverse")
library("DESeq2")
```

Trying to find non diff expr genes
So in terms of: Failure to reject the null Hypothesis cannot be taken as evidence of no difference!
```{r}
LAKE <- "LM"
LAKE1 <- "LT"
FILE.DIR     <- paste("directory")
FILE.PATH <- paste(FILE.DIR, "/Output/", LAKE, "/", sep = "")
FILE.PATH1 <- paste(FILE.DIR, "/Output/", LAKE1, "/", sep = "")
PATH.OUT <- paste(FILE.DIR, "/Output/nondiff_refgene_search/", sep = "")
DATE <- format(Sys.time(), "%m%d%Y")

# set directory 
setwd(FILE.DIR)

LM <- read_csv(file = paste0(FILE.PATH, LAKE, "_ordered_pvalue_005.csv"))
LMO <- LM[order(LM$padj),]

LMO_2fold_low <- subset(LMO, log2FoldChange > -0.05)
nrow(LMO_2fold_low)
LMO_2fold_final <- subset(LMO_2fold_low, log2FoldChange < 0.05)
nrow(LMO_2fold_final)
LMO_2fold <- LMO_2fold_final[order(LMO_2fold_final$padj),]
write.csv(as.data.frame(LMO_2fold), file = paste(PATH.OUT, LAKE, "_ordered_nondiff", "_", DATE, ".csv", sep = ""))

lmf <- LMO_2fold
# separate gene_id and ensembl_id
lmf_names <- lmf %>%
  dplyr::select(X1) %>% 
  separate(X1,
           c("gene_id", "ensembl"),
           sep = "[|]", remove = FALSE)
# replace ensembl_ids that are missing with gene_id
lmf_names$ensembl <- ifelse(is.na(lmf_names$ensembl), lmf_names$gene_id, lmf_names$ensembl)
# merge names and data frame
lmfo <- merge(lmf_names, lmf,  by = "X1")
#lmfo <- lmfo[order(lmfo$baseMean, decreasing = TRUE),]
nrow(lmfo)


LT <- read_csv(file = paste0(FILE.PATH1, LAKE1, "_ordered_pvalue_005.csv"))
LTO <- LT[order(LT$padj),]

LTO_2fold_low <- subset(LTO, log2FoldChange > -0.05)
nrow(LTO_2fold_low)
LTO_2fold_final <- subset(LTO_2fold_low, log2FoldChange < 0.05)
nrow(LTO_2fold_final)
LTO_2fold <- LTO_2fold_final[order(LTO_2fold_final$padj),]
write.csv(as.data.frame(LTO_2fold), file = paste(PATH.OUT , LAKE1, "_ordered_nondiff", "_", DATE, ".csv", sep = ""))

ltf <- LTO_2fold
nrow(ltf)
ltf_names <- ltf %>%
  dplyr::select(X1) %>% 
  separate(X1,
           c("gene_id", "ensembl"),
           sep = "[|]", remove = FALSE)
# replace ensembl_ids that are missing with gene_id
ltf_names$ensembl <- ifelse(is.na(ltf_names$ensembl), ltf_names$gene_id, ltf_names$ensembl)
# merge names and data frame
ltfo <- merge(ltf_names, ltf,  by = "X1")
#ltfo <- ltfo[order(ltfo$baseMean, decreasing = TRUE),]
nrow(ltfo)

```




Find lowest variation across samples
```{r}

# all <- read_csv(file = paste(PATH.OUT,LAKE,"_", LAKE1,"_overlap_ordered_nondiff", "_", DATE, ".csv", sep = ""))

LMnorm <- read_csv(file = paste0(FILE.DIR,"/normalized_gcmatrix_", LAKE, ".csv"))
nrow(LMnorm)
LTnorm <- read_csv(file = paste0(FILE.DIR,"/normalized_gcmatrix_", LAKE1, ".csv"))
nrow(LMnorm)


# norm <- LMnorm %>% 
#   left_join(., LTnorm, by = c("X1"))

norm_name_LM <- LMnorm %>% 
  dplyr::select(X1) %>% 
  separate(X1,
           c("gene_id", "ensembl"),
           sep = "[|]", remove = FALSE)
# replace ensembl_ids that are missing with gene_id
norm_name_LM$ensembl <- ifelse(is.na(norm_name_LM$ensembl), norm_name_LM$gene_id, norm_name_LM$ensembl)
normLM <- merge(norm_name_LM, LMnorm,  by = "X1")

norm_name_LT <- LTnorm %>% 
  dplyr::select(X1) %>% 
  separate(X1,
           c("gene_id", "ensembl"),
           sep = "[|]", remove = FALSE)
# replace ensembl_ids that are missing with gene_id
norm_name_LT$ensembl <- ifelse(is.na(norm_name_LT$ensembl), norm_name_LT$gene_id, norm_name_LT$ensembl)
normLT <- merge(norm_name_LT, LTnorm,  by = "X1")

subLM <-  normLM$X1 %in% lmfo$X1
genes_LM <- lmfo[subLM,]

subLT <-  normLT$X1 %in% ltfo$X1
genes_LT <- ltfo[subLT,]

LM_cov <- normLM %>% 
  mutate(vars = rowVars(as.matrix(normLM[4:13]))) %>% 
  mutate(cov = rowSds(as.matrix(normLM[4:13]),na.rm = TRUE)/rowMeans(as.matrix(normLM[4:13]))*100)

LT_cov <- normLT %>% 
  mutate(vars = rowVars(as.matrix(normLT[4:13]))) %>% 
  mutate(cov = rowSds(as.matrix(normLT[4:13]),na.rm = TRUE)/rowMeans(as.matrix(normLT[4:13]))*100)

#norm_var1 <- norm_var[order(norm_var$cov, decreasing = FALSE),]

addonLM <- LM_cov %>% 
  dplyr::select(X1, vars, cov)

LMfo <- lmfo %>% 
  left_join(addonLM, all, by = c("X1" = "X1"))

LMfo <- LMfo[order(LMfo$cov, decreasing = FALSE),]


addonLT <- LT_cov %>% 
  dplyr::select(X1, vars, cov)

LTfo <- ltfo %>% 
  left_join(addonLT, all, by = c("X1" = "X1"))

LTfo <- LTfo[order(LTfo$cov, decreasing = FALSE),]

  
all <- LTfo %>% 
  inner_join(LMfo, by = "X1", suffix = c("LT", "LM"))
nrow(all)
write.csv(all, file = paste(PATH.OUT,LAKE,"_", LAKE1,"_ref_genes_ordered_", DATE, ".csv", sep = ""), row.names = FALSE)
  

allnose <- all %>% 
  dplyr::select(X1, ensemblLT, baseMeanLT, log2FoldChangeLT, padjLT, varsLT, covLT, baseMeanLM, log2FoldChangeLM, padjLM, varsLM, covLM)

fin <- dplyr::rename(allnose, ensembl = ensemblLT)

overlap_ensembl <- fin %>% 
  dplyr::select(
    ensembl
  )

nrow(overlap_ensembl)
write.csv(overlap_ensembl, file = paste(PATH.OUT,LAKE,"_", LAKE1,"_overlap_ordered_nondiff_ensemblID", "_", DATE, ".csv", sep = ""), row.names = FALSE)

#select 
summary(fin$covLT)
summary(fin$covLM)

# not sure if it makes sense. Maybe go only with the single covLT and covLM
fin <- fin %>% 
  mutate(
    cov = covLT + covLM
  ) %>% 
  arrange(cov) %>% 
  dplyr::select(- varsLT, - varsLM, -padjLT, - padjLM)

fin25 <- fin[fin$cov < quantile(fin$cov , 0.25),]

godata <- read_csv(file = paste0(FILE.DIR, "/Output/LM_biomaRt_GO_id.csv"))

external_name <- godata %>% 
  dplyr::select(1,2) %>%
  unique()

ref_genes <- fin25 %>% 
  left_join(., external_name, by =c("ensembl" = "ensembl_gene_id")) %>% 
  drop_na()

nrow(ref_genes)
write_csv(ref_genes, file = paste(PATH.OUT,LAKE,"_", LAKE1,"_smallCV_", DATE, ".csv", sep = ""))
```




